Stack Safety as a Security Property

### 2021-02-15 main.tex
Add context to states

We next describe our machine model. To make our definition of stack safety
as generic as possible, this section proposes an abstract interface to a
machine model plus some kind of
% static or dynamic,
(software- or
hardware-enforced) policy monitor. In~\cref{sec:enforcement}, we will
instantiate this model with a concrete dynamic policy enforcement mechanism
based on \citet{DBLP:conf/sp/RoesslerD18}.

@@

The basic building blocks of the machine are {\em values} and {\em
  addresses}. Both are drawn from some set of {\em words}
$\WORDS$, ranged over by $\word$.
%
Our machine states are composed of {\em components} $k$, addressed either by
words (memory addresses) or by register names ($r$) drawn from some set
$\REGS$.
%
    \[\component \in \COMPONENTS = \WORDS + \REGS \]
%
The register names are assumed to include two special purpose registers: the
program counter {\PCname} and the stack pointer \SP.

@@

The step function of the machine takes a machine state and produces a
pair of a machine state and an {\em observation}, which may be either
silent (written $\tau$) or a word of data\ifaftersubmission\bcp{This is a rather concrete
  choice!  Why not leave the set of observations abstract?}\fi that the machine
communicates to the outside world:
%
\[\obs \in \OBSS = \{\tau\} \uplus \WORDS\]
\[\mach \stepstoobs{\obs} \mach' \in \MACHS \rightarrow \MACHS \times \OBSS \]
%
This step function is total, so the machine's execution traces
(defined in \cref{sec:traces}) are infinite.

@@

We now present our model of enforcement, which we introduced
informally in \Cref{sec:threat}.  A policy is a general model of an
enforcement mechanism, consisting of a set of policy states \(\pol \in
\POLS\), a policy step function \((\mach, \pol) \polstep \pol' \in
\MACHS \times \POLS \rightharpoonup \POLS\). This step function is partial;
it is undefined on input configurations that correspond to a policy fault.
We assume that the
compiler provides an initial policy state $\pol_0$ to accompany
the initial machine state.
%
A concrete policy based on \citet{DBLP:conf/sp/RoesslerD18} will be described in
\Cref{sec:enforcement}.

@@

@@

A trace over a type $\TRACEELEMS$ is a nonempty, finite or infinite sequence of
$\TRACEELEMS$s; we write $\TRACE{\TRACEELEMS}$ for the set of traces over
$\TRACEELEMS$. We will mostly be concerned with sequences of pairs of
machine states
and observations, called \emph{machine traces} and denoted $\machT \in \MOTRACE$,
and with sequences of triples of a machine state, a policy state, and an
observation, called \emph{MP-traces} and denoted $\MPT \in \MPOTRACE$.

@@

We will frequently take a prefix of the trace from an initial state up
through the first state where some condition holds on the machine state.
This can easily be implemented using \(\PUT\), and we provide special notation
for convenience.
%
If \(f\) be a predicate on machine states, we define \(\mach
\hookrightarrow \machT | f\) (read ``\(\machT\) is the prefix of
running \(\mach\) up to \(f\)'') and similarly for traces with policies:

>>>>

We next describe our machine model. To make our definition of stack safety
as generic as possible, this section proposes an abstract interface to a
machine model plus some kind of (software- or hardware-enforced) policy monitor.
In~\cref{sec:enforcement}, we will instantiate this model with a concrete dynamic
policy enforcement mechanism based on \citet{DBLP:conf/sp/RoesslerD18}.

@@

The basic building blocks of the machine are {\em values} and {\em
  addresses}. Both are drawn from some set of {\em words}
\(\WORDS\), ranged over by \(\word\).
%
Our machine states are composed of {\em components} \(\component\), addressed either by
words (memory addresses) or by register names (\(\reg\)) drawn from some set
\(\REGS\).
%
    \[\component \in \COMPONENTS = \WORDS + \REGS \]
%
The register names are assumed to include two special purpose registers: the
program counter {\PCname} and the stack pointer \(\SP\).

@@

The step function of the machine takes a machine state and produces a
pair of a machine state and an {\em observation}, which may be either
silent (written \(\tau\)) or a word of data\ifaftersubmission\bcp{This is a rather concrete
  choice!  Why not leave the set of observations abstract?}\fi that the machine
communicates to the outside world:
%
\[\obs \in \OBSS = \{\tau\} \uplus \WORDS\]
\[\mach \stepstoobs{\obs} \mach' \in \MACHS \rightarrow \MACHS \times \OBSS \]
%
This step function is total, so the machine's execution traces
(defined in \cref{sec:traces}) are infinite.

@@

We now present our model of enforcement, which we introduced informally in \Cref{sec:threat}.
A policy is a general model of an enforcement mechanism, consisting of a set of policy
states \(\pol \in \POLS\) and a policy step function \((\mach, \pol) \polstep \pol' \in
\MACHS \times \POLS \rightharpoonup \POLS\). This step function is partial;
it is undefined on input configurations that correspond to a policy fault. The policy
can see the machine state, but cannot modify it. We assume that the
compiler provides an initial policy state \(\pol_0\) to accompany the initial machine state.
%
A concrete policy based on \citet{DBLP:conf/sp/RoesslerD18} will be described in
\Cref{sec:enforcement}.

@@

\subsection{Contexts}

Separate from the machine state, our formal properties will keep track of additional context
information representing the history of the run. The following definitions are parameterized
over an arbitrary set of context states, \(\CONTEXTS\), an initial context state \(\context_0
\in \CONTEXTS\), and a context step function \(\mach, \context \constep \context' \in
\MACHS \times \CONTEXTS \rightarrow \CONTEXTS\).

We lift contexts into an MP trace in the same way as we lifted policy states to get
a state type \(\mpcstatename \in \MPCS = \MACHS \times \POLS \times \CONTEXTS\).

\judgmenttwo{\(\mpstate{\mach_1}{\pol_1} \stepstopol{\obs} \mpstate{\mach_2}{\pol_2}\)}
            {\(\mach_1, \context_1 \constep \context_2\)}
            {\(\mpcstate{\mach_1}{\pol_1}{\context_1} \stepstocon{\obs}
               \mpcstate{\mach_2}{\pol_2}{\context_2}\)}

@@

A trace over a type \(\TRACEELEMS\) is a nonempty, finite or infinite sequence of
\(\TRACEELEMS\)s; we write \(\TRACE{\TRACEELEMS}\) for the set of traces over
\(\TRACEELEMS\). We will mostly be concerned with sequences of pairs of machine states
and observations, called \emph{machine traces} and denoted \(\machT \in \MOTRACE\),
and with sequences of 4-tuples of a machine state, a policy state, a context, and an
observation, called \emph{MPC-traces} and denoted \(\MPCT \in \MPCOTRACE\).

@@

We will frequently take a prefix of the trace from an initial state up
through the first state where some condition holds on the machine state.
This can easily be implemented using \(\PUT\), and we provide special notation
for convenience.
%
If \(f\) a predicate on machine states, we define \(\mach
\hookrightarrow \machT | f\) (read ``\(\machT\) is the prefix of
running \(\mach\) up to \(f\)'') and similarly for traces with policies:
